---
title: "DATA 606 Data Project Proposal"
author: "KoohPy <- Koohyar Pooladvand"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document:
    latex_engine: lualatex  # Use lualatex engine instead of pdf_engine
  word_document: 
editor_options:
  markdown:
    wrap: 72
bibliography: references.bib
---

### Data Preparation

In this project, I have chosen to work on breast cancer. There are
various resources available regarding this particular topic, with the
SEER being the most reliable one.

The Surveillance, Epidemiology, and End Results (SEER) Program of the
National Cancer Institute (NCI) collects and publishes cancer data
through a coordinated system of strategically placed cancer registries,
which cover nearly 30% of the US population.

Currently, there are 18 SEER registries in the USA. This information can
be found on the following website:
<https://seer.cancer.gov/data/access.html>.

I have also used the following repository to assist me with this
project: <https://github.com/kohyarp/SEER_solid_tumor>. The Database
contains tons of data, the goal of my investigation will be focused only
on BREAST cancer for 2011-2015 and 2019-2020. SEER has a software \*STAT
that I have used to import the data to a test that will be stored and
used on my local computer. Additionally there is a GITHUB repository
that I have used to some extent in this project. The repository is
focused on all type of cancer, but my study is focused on BREAST, and I
aim different question to answer.
<https://github.com/zgalochkina/SEER_solid_tumor>

```{r Code_initialization, echo= FALSE, results='hide', warning=FALSE, message=FALSE}
required_packages <- c("RSQLite","devtools","tidyverse","DBI","dplyr","odbc","openintro","ggplot2","psych","reshape2","knitr","markdown","shiny","R.rsp","fivethirtyeight","RCurl", "stringr","readr","glue","data.table", "hflights", "jsonlite", "rjson", "XML", "xml2", "rvest", "readxl", "openxlsx", "httr2","kableExtra", "tinytex", "tidytext","textdata", "wordcloud", "plyr", "survival", "scales", "stats", "forcats", "gridExtra", "plotrix", "maptools", "mapproj", "rgeos", "rgdal", "rmarkdown", "R.utils", "utils", "skimr","GGally")

not_installed <- required_packages[!(required_packages %in% installed.packages()[ , "Package"])]

if(length(not_installed) == 0) {
  print("All required packages are installed")
} else {
  print(paste(length(not_installed), "package(s) need to be installed.")) # Print the list of packages that need to be installed
  tryCatch({
    install.packages(not_installed, dependencies = TRUE)
  }, error = function(e) {
    cat("Error occurred during package installation:", e$message, "\n")
  })
}


#remotes::install_github("cran/SEER2R")
#remotes::install_github("cran/maptools")

# define different paths to load the files 
library(kableExtra)
library(tinytex)
library(textdata)
library(tidytext)
library(tidyr)
library(ggplot2)
library(wordcloud)
library(dplyr)
library(reshape2)
library(plyr)       # for 'ddply' function
library(tidyverse)  # filter, transform, plot data with 'ggplot2', 'tibble', 'tidyr', 'readr', 'purrr', 'dplyr' packages
library(SEER2R)     # read SEER data files
library(survival)   # Kaplan-Meier, Cox
library(scales)     # transform data scale on plot
library(stats)      # for 'fisher.test', 'aov' functions
library(stringr)    # for 'str_replace_all' function
library(forcats)    # for 'fct_recode' function
library(gridExtra)  # for 'grid.arrange' function
library(plotrix)    # for 'addtable2plot' function
#library(maptools)   # map data to 50 states
library(mapproj)    # map data to 50 states
#library(rgeos)      # map data to 50 states
#library(rgdal)      # map data to 50 states
library(knitr)      # for reports
library(R.utils)    # to unzip .gz population data file
library(utils)      # to unzip .zip GIS files for maps
library(skimr)

#surpass the error message for dplyr to not show the masking
suppressPackageStartupMessages(library(dplyr))

```

### Research question

The primary question I aim to address is the survival rate of breast
cancers and the influence of factors such as age, type, sex, and other
parameters on this rate. Notably, a five-year threshold is commonly used
to determine survival rates. Although my understanding of the rationale
behind this five-year benchmark is limited, recognizing its significance
has led me to divide the data into two separate datasets.

The dataset spanning from 2011-2015 assumes that the status of all
patients within that period is known up to the database’s current date
in 2022. Additionally, I have selected the most recent data from
2019-2020 as my target years for potential correlation and regression
studies to estimate survival rates.

This analysis is not scientific but rather a straightforward statistical
exercise with no purpose beyond this course. However, I find the subject
intriguing to investigate. I am uncertain if I will discover any
significant relationships or correlations, and if found, whether they
will be relevant, as I am not an expert in the field of breast cancer.
My choice of topic is personal, as I have witnessed immediate family
members diagnosed with this cancer, and I wish to gain a deeper
understanding.

The database for 2011-2015 contains approximately 303,000 rows with 36
selected columns. I have chosen to focus solely on the 2019-2020 data,
which comprises about 131,000 rows for prediction purposes. The question
at hand is complex, and while I do not anticipate a definitive answer, I
hope to uncover some patterns and test hypotheses, as well as engage in
general data work, from tidying to cleaning.

Furthermore, I plan to explore regression analysis to determine if I can
identify any linear or non-linear relationships among the critical
parameters.

My knowledge of the subject is not extensive, but I am eager to learn as
I progress.

Some of the general parameters to consider are as follows: \* Years of
diagnoses; \* Age groups at diagnosis; \* Cancer type (BREAST);

Some other parameters are also available to be edited, but they are
secondary.

"to be added : adding a brief literature review to provide context for
my research questions and hypotheses. This could include previous
studies on breast cancer survival rates, factors affecting survival, and
methods used for analysis."

```{r, import_data, echo = TRUE}

# Replace "file.txt" with the path to your text file
directory <- "C:/Users/kohya/OneDrive/CUNY/DATA 606/DATA 606 Spring/Project"
file_2020 <- "BREAST_2019-2020-updated.csv"
file_serv <- "BREAST_2011-2015.csv"
# Complete the file path
full_path_serv <- file.path(directory, file_serv)
full_path_eval<- file.path(directory, file_2020)


BREAST_DF_surv <- read.csv(full_path_serv, header = TRUE,
                      na.strings = "NA", check.names = FALSE)
BREAST_DF_eval <- read.csv(full_path_eval, header = TRUE,
                      na.strings = "NA", check.names = FALSE)

labels_of_interest <- c("Primary Site - labeled")

# View the first few rows of the data frame
kable(head(BREAST_DF_surv, 10))

kable(head(BREAST_DF_eval, 10))

```

### Cases

**What are the cases, and how many are there?** There are 131,395 cases
in the BREAST cancer list of 2019-2020. And There are 303557 in
2011-2015 dataset.

"adding more exploratory data analysis (EDA) to understand the structure
and distribution of variables in your dataset. This could include
summary statistics, histograms, scatter plots, or other visualizations."

By employing Exploratory Data Analysis (EDA) methods like summary
statistics and graphical representations, we aim to reveal insights that
will enhance our comprehension of breast cancer outcomes and therapeutic
approaches. The dataset is rich with details, encompassing variables
such as the patient’s age at operation, operation year, count of
positive axillary nodes detected, and survival status post-treatment.

[https://medium.com/\@navamisunil174/exploratory-data-analysis-of-breast-cancer-survival-prediction-dataset-c423e4137e38](https://medium.com/@navamisunil174/exploratory-data-analysis-of-breast-cancer-survival-prediction-dataset-c423e4137e38){.uri}

### Data collection

**Describe the method of data collection.** I used the SEER \*STAT to
collect the data and export it as a TXT to be able to import it to the R
for analyses. How SEER collects the data is explained in the following
page in summary:

-   The SEER program collects cancer incidence data through a network of
    population-based cancer registries. These registries gather
    information on patient demographics, primary tumor site, tumor
    morphology, stage at diagnosis, and first course of treatment. They
    also follow up with patients for vital status.

-   By law, these facilities are required to report new cancer cases to
    a central cancer registry, like a state cancer registry.

-   [The SEER program releases new research data annually, based on
    submissions from the previous year, and makes it available for
    public use through a data request
    process](https://seer.cancer.gov/data-software/). This comprehensive
    approach ensures that the SEER database is a valuable resource for
    cancer research and surveillance.

<https://training.seer.cancer.gov/registration/data/collection.html>

### Type of study

This will be an observational study, information is gathered for
different patients and I will be evaluating the available data to
present and evaluate.

"discussing potential limitations of observational studies, such as
confounding variables and biases, and how you plan to address them in
analysis."

**What type of study is this (observational/experiment)?**

### Data Source

Data is collected from SEER program and I used SEER \*STAT software to
glean them in a format that can be used and imported as TXT to R
[@SEER2023].

"providing additional details about the specific variables included in
dataset and how they were collected"

**If you collected the data, state self-collected. If not, provide a
citation/link.**

### Dependent Variable

I am still looking into the data, it seems I will have a combination of
both quantitative and qualitative data to work with. For example, while
the number of tumors, and survival months are qualitative. Other like
race, marital status, type of cancer are categorical. I am still looking
to see if I can find any qualitative data.

Categorical features, such as ‘Median household income ...’ ‘Marital
Status,’ ‘Grade recode’ ‘laterality’ and ‘Radiatio recode' and so on are
represented as objects (characters).

Integer data types (int64) are assigned to ‘Patient ID,’ ‘Year of
diagnosis,’ ‘total number of ...'.

*The event indicator refers to the death and the time registered is
either the time-to-event (when the individual eventually dies) or the
time-to-censorship (the event is not observed), measured in months.*

```{r variable_structure, echo=TRUE}

# Find unique values in each column
# Apply function to find unique values for each column

unique_values <- data.frame(unique = apply(BREAST_DF_surv, 2, function(x) length(unique(x))),colnames = colnames(BREAST_DF_surv))

# Check for NULL values
any_null <- any(sapply(BREAST_DF_surv, is.null))

# Check for NA values
any_na <- any(sapply(BREAST_DF_surv, is.na))

# Check if there are any NULL or NA values
if (any_null || any_na) {
  print("The data frame contains NULL or NA values.")
} else {
  print("The data frame does not contain any NULL or NA values.")
}

has_na_character <- any(sapply(BREAST_DF_surv, function(x) any(x == "NA")))

if (has_na_character) {
  print("The data frame contains character values of 'NA'.")
} else {
  print("The data frame does not contain character values of 'NA'.")
}

```

### Data tyding 

Upon exploring the data, it seems data might have an empty column, in
this data-based, the empty values are filled with "Blanks". Thus, in
this section, I first explore if there is any column which is entirely
empty, then will remove it and if there are others which have some empty
values filled with "blancked" I will repalced them with "NA" whoch is
handled better in dplyr and tydiverse.

```{r data_tyding, echo=TRUE}

# There are cells in the DF that contianes "Blank(s) which is literally NA, first I want to find if there is any column that all is values is Blank(s), if then remove them.

#look for columns with all "Blank(s)" values
Empty_column <- BREAST_DF_surv %>%
  dplyr::summarise(dplyr::across(everything(), ~all(. == "Blank(s)"))) %>%
  as.logical() %>%
  unlist()

# Get the names of columns with all cells containing "Blank(s)"
blank_column_names <- names( BREAST_DF_surv)[Empty_column]

# Print the column names with all cells containing "Blanks"
print(blank_column_names)

#remove those empty column from thr DF
BREAST_DF_surv <- BREAST_DF_surv[, !names(BREAST_DF_surv) %in% blank_column_names]
BREAST_DF_eval <- BREAST_DF_eval[, !names(BREAST_DF_eval) %in% blank_column_names]

#Then let's see if there is any cell in the remaining that migth still have "Blank(s)", if so repalce it with NS which is betetr handle in R

#This code first replaces all occurrences of "Blank(s)" with an empty string "", and then uses na_if() to convert the empty strings to NA. Now, all cells that previously had "Blank(s)" are replaced with NA, making it easier to handle missing values in R.

BREAST_DF_surv <- BREAST_DF_surv %>%
  mutate_if(is.character, ~ifelse(. == "Blank(s)", "", .)) %>%  # For character columns
  mutate_if(is.numeric, ~ifelse(. == "", as.numeric(NA), .))  # For numeric columns

# Now, empty character cells are replaced with NA
BREAST_DF_surv <- BREAST_DF_surv %>%
  mutate_if(is.character, na_if, "")


#same to be done for eval dataset


BREAST_DF_eval <- BREAST_DF_eval %>%
  mutate_if(is.character, ~ifelse(. == "Blank(s)", "", .)) %>%  # For character columns
  mutate_if(is.numeric, ~ifelse(. == "", as.numeric(NA), .))  # For numeric columns

# Now, empty character cells are replaced with NA
BREAST_DF_eval <- BREAST_DF_eval %>%
  mutate_if(is.character, na_if, "")

#Change characters to numerics 
BREAST_DF_surv$`Months from diagnosis to treatment` <- as.numeric(BREAST_DF_surv$`Months from diagnosis to treatment`)
BREAST_DF_surv$`Survival months` <- as.numeric(BREAST_DF_surv$`Survival months`)
BREAST_DF_surv$`Total number of in situ/malignant tumors for patient` <- 
  as.numeric(BREAST_DF_surv$`Total number of in situ/malignant tumors for patient`)
BREAST_DF_surv$`Total number of benign/borderline tumors for patient` <- 
  as.numeric(BREAST_DF_surv$`Total number of benign/borderline tumors for patient`)
#Change the character to numeric in Eval dataset too
BREAST_DF_eval$`Months from diagnosis to treatment` <- as.numeric(BREAST_DF_eval$`Months from diagnosis to treatment`)
BREAST_DF_eval$`Survival months` <- as.numeric(BREAST_DF_eval$`Survival months`)
BREAST_DF_eval$`Total number of in situ/malignant tumors for patient` <- 
  as.numeric(BREAST_DF_eval$`Total number of in situ/malignant tumors for patient`)
BREAST_DF_eval$`Total number of benign/borderline tumors for patient` <- 
  as.numeric(BREAST_DF_eval$`Total number of benign/borderline tumors for patient`)


# View the structure of the data frame
#str(BREAST_DF_surv)
skimr::skim(BREAST_DF_surv)
skimr::skim(BREAST_DF_eval)

```

**What is the response variable? Is it quantitative or qualitative?**

### Independent Variable(s)

### Relevant summary statistics

**Provide summary statistics for each the variables. Also include
appropriate visualizations related to your research question (e.g.
scatter plot, boxplots, etc). This step requires the use of R, hence a
code chunk is provided below. Insert more code chunks as needed.**

```{r summary_statistic_1, echo=TRUE }

#find column name to use later if needed
DF_col_names <- colnames(BREAST_DF_surv)

#Find unique values in `Race recode (W, B, AI, API)` column
uniques_races <- unique(BREAST_DF_surv$`Race recode (W, B, AI, API)`)

# use ggplot to plot the race information 
BREAST_DF_surv |> 
  ggplot(mapping = aes(x=`Race recode (W, B, AI, API)`)) +
  geom_bar(stat = "count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(label = after_stat(count)), stat = "count", vjust = -0.5) +
  ylim(0, 246000)

#we want to coampre the percentage of the diferent race in the eval and survival data, thus i use sumamrise to create two new DF to only store the sumamry statistic specifically including the percentage of race based on the population
#find percentage of race for the survival
BREAST_DF_perc_surv <- BREAST_DF_surv %>%
  group_by(`Race recode (W, B, AI, API)`) %>%
  dplyr::summarise(count = dplyr::n()) %>%  # Calculate count per group
  ungroup() %>%  # Ungroup the data
  mutate(total_count = sum(count)) %>%  # Calculate total count
  mutate(percentage = count / total_count * 100)  # Calculate percentage using total count

# Plot the percentages
ggplot(BREAST_DF_perc_surv, aes(x = `Race recode (W, B, AI, API)`, y = percentage)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), vjust = -0.5, color = "black") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Percentage of Population by Race between 2011-2015", x = "Race recode (W, B, AI, API)", y = "Percentage") + ylim (0,90)

BREAST_DF_eval |> 
  ggplot(mapping = aes(x=`Race recode (W, B, AI, API)`)) +
  geom_bar(stat = "count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(label = after_stat(count)), stat = "count", vjust = -0.5) +
  ylim(0, 104000)

BREAST_DF_perc_eval <- BREAST_DF_eval %>%
  group_by(`Race recode (W, B, AI, API)`) %>%
  dplyr::summarise(count = dplyr::n()) %>%  # Calculate count per group
  ungroup() %>%  # Ungroup the data
  mutate(total_count = sum(count)) %>%  # Calculate total count
  mutate(percentage = count / total_count * 100)  # Calculate percentage using total count

# Plot the percentages
ggplot(BREAST_DF_perc_eval, aes(x = `Race recode (W, B, AI, API)`, y = percentage)) +
  geom_bar(stat = "identity", fill = "plum") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), vjust = -0.5, color = "black") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Percentage of Population by  between 2019-2022", x = "Race recode (W, B, AI, API)", y = "Percentage") + ylim (0,90)



# In this section I want to focus on the age and see if age matetrs, same sets of data is going to be plot for ages, starting with percentage for eval and surve 
#find percentage of race for the survival
#find ubique values for column ratted to age 
uniques_ages <- unique(BREAST_DF_surv[29])

BREAST_DF_age_perc_surv <- BREAST_DF_surv %>%
  dplyr::group_by(`Age recode (<60,60-69,70+)`) %>%
  dplyr::summarise(count = dplyr::n()) %>%  # Calculate count per group
  ungroup() %>%  # Ungroup the data
  mutate(total_count = sum(count)) %>%  # Calculate total count
  mutate(percentage = count / total_count * 100)  # Calculate percentage using total count

perc_max <- max(BREAST_DF_age_perc_surv$percentage)
# Plot the percentages
ggplot(BREAST_DF_age_perc_surv, aes(x = `Age recode (<60,60-69,70+)`, y = percentage)) +
  geom_bar(stat = "identity", fill = "brown") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), hjust = -0.1 , vjust = 0.4, color = "black", angle = 90) +  # Rotate the text vertically
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +labs(title = "Percentage of Population by Age range 2011-2015", 
       x = "Age range", 
       y = "Percentage") + 
  ylim(0, round(1.5 * perc_max, 1))

# In this section we do the same analyses for Eval dta based on age
BREAST_DF_age_perc_eval <- BREAST_DF_eval %>%
  dplyr::group_by(`Age recode (<60,60-69,70+)`) %>%
  dplyr::summarise(count = dplyr::n()) %>%  # Calculate count per group
  ungroup() %>%  # Ungroup the data
  mutate(total_count = sum(count)) %>%  # Calculate total count
  mutate(percentage = count / total_count * 100)  # Calculate percentage using total count

# Plot the percentages
ggplot(BREAST_DF_age_perc_eval, aes(x = `Age recode (<60,60-69,70+)`, y = percentage)) +
  geom_bar(stat = "identity", fill = "brown") +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), hjust = -0.1 , vjust = 0.4, color = "black", angle = 90) +  # Rotate the text vertically
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +labs(title = "Percentage of Population by Age range 2019-2022", 
       x = "Age range", 
       y = "Percentage") + 
  ylim(0, round(1.5 * perc_max, 1))


# In this section, we do the analyses on household income} 
#find ubique values for column ratted to age 
uniques_householdes <- unique(BREAST_DF_surv[27])

BREAST_DF_income_perc_surv <- BREAST_DF_surv %>% dplyr::group_by(`Median household income inflation adj to 2021`) %>% 
  dplyr::summarise(count = dplyr::n()) %>% # Calculate count per group 
  ungroup() %>% # Ungroup the data 
  mutate(total_count = sum(count)) %>% # Calculate total count 
  mutate(percentage = count / total_count * 100) # Calculate percentage using total count

perc_max <- max(BREAST_DF_income_perc_surv$percentage) # Plot the percentages 
ggplot(BREAST_DF_income_perc_surv, aes(x = `Median household income inflation adj to 2021`, y = percentage)) + 
  geom_bar(stat = "identity", fill = "brown") + 
  geom_text(aes(label = paste0(round(percentage, 1), "%")), hjust = -0.1 , vjust = 0.4, color = "black", angle = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Percentage of Population by income 2011-2015", x = "Household Income", y = "Percentage") + 
  ylim(0, 1.2*perc_max)


#In this section we do the same analyses for Eval data based on age
BREAST_DF_income_perc_eval <- BREAST_DF_eval %>% 
  dplyr::group_by(`Median household income inflation adj to 2021`) %>% 
  dplyr::summarise(count = dplyr::n()) %>% # Calculate count per group 
  ungroup() %>% # Ungroup the data 
  mutate(total_count = sum(count)) %>% # Calculate total count 
  mutate(percentage = count / total_count * 100) # Calculate percentage using total count


#Plot the percentages
perc_max <- max(BREAST_DF_income_perc_eval$percentage)
ggplot(BREAST_DF_income_perc_eval, aes(x = `Median household income inflation adj to 2021`, y = percentage)) + 
  geom_bar(stat = "identity", fill = "brown") + 
  geom_text(aes(label = paste0(round(percentage, 1), "%")), hjust = -0.1 , vjust = 0.4, color = "black", angle = 0) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Percentage of Population by income 2019-2022", x = "Household Income", y = "Percentage") + 
  ylim(0, 1.2*perc_max)



```

```{r more_sumamry, echo=TRUE}

# In this section I want to focus on the cause of dead, COD, and investigate whether those who have had cancer are alive, anf if no what was the cause of dead. 
#find percentage of deceased due to breast cancer
#find unique values for column ratted to age 

uniques_CODs <- unique(BREAST_DF_surv[20])
DF_col_names[20]

# check if the column `COD to site recode` has value of Alive or Breast meaning they are still alive or have died because of breast cancer, and other passed a way but not because of Breast cancer. 

BREAST_DF_surv <- BREAST_DF_surv %>%
  mutate(COD = ifelse(`COD to site recode` %in% c("Alive","Breast"), `COD to site recode`, "Other"))


BREAST_DF_COD_perc_surv <- BREAST_DF_surv %>%
  dplyr::group_by(COD) %>%
  dplyr::summarise(count = dplyr::n()) %>%  # Calculate count per group
  ungroup() %>%  # Ungroup the data
  mutate(`Total Count` = sum(count)) %>%  # Calculate total count
  mutate(Population = round(count / `Total Count` * 100),2)  # Calculate percentage using total count

kable(BREAST_DF_COD_perc_surv)

# Let's first group by the number of tumor and find hom many in the population have those and then among them look how many passed away only due to breast. It is not completely correct, becuse thre are some that migth have passed away due to Breast cancer complication that is not in this counts. 
BREAST_DF_TNoT_perc_surv <- BREAST_DF_surv %>%
  dplyr::group_by(`Total number of in situ/malignant tumors for patient`) %>%
  dplyr::add_count() %>%
  filter(COD == "Breast") %>%
  dplyr::summarise(`Event Population` = n(), 
            Population = dplyr::first(n))  # Use `first()` to extract the total count in each 

# Do simple math to fidn the percentage of the groupn un the population and then the percentage of the deceased within the group. 

BREAST_DF_TNoT_perc_surv$`Group % in total` <- round(BREAST_DF_TNoT_perc_surv$Population/sum(BREAST_DF_TNoT_perc_surv$Population)*100,2)

BREAST_DF_TNoT_perc_surv$`Death %` <- round(BREAST_DF_TNoT_perc_surv$`Event Population`/BREAST_DF_TNoT_perc_surv$Population*100,2)

    
kable(BREAST_DF_TNoT_perc_surv)


# Let' focus on the treatemnt, There are two type of treatment and can be a 4 combination ,as follows: Radiation: R, Chemoteraphy: C,  R:N-C:N,  R:Y-C:N, R:N-C:Y, R:Y-C:Y. We must look into these 4 group and find the total number and then in each find the number of death. Finally report them imialrly that we have done above. 

BREAST_DF_surv <- BREAST_DF_surv %>% 
  mutate(Radiation = ifelse(`Radiation recode` %in% c("None/Unknown","Refused (1988+)","Recommended, unknown if administered"),"No/Unknown","Yes"))

#use DPLYR to filter based on two parameters chemotheraphy and radiation therapy and evalaute the death rate accordingly  
BREAST_DF_RNC_perc_surv <- BREAST_DF_surv %>%
  dplyr::group_by(Radiation,`Chemotherapy recode (yes, no/unk)`) %>%
  dplyr::add_count() %>%
  filter(COD == "Breast") %>%
  dplyr::summarise(`Event Population` = n(), 
            Population = dplyr::first(n))  # Use `first()` to extract the total count in each 

#knwoign the population calcualte the gorup rate and death rate in each group 
BREAST_DF_RNC_perc_surv$`Group % in total` <- round(BREAST_DF_RNC_perc_surv$Population/sum(BREAST_DF_RNC_perc_surv$Population)*100,2)

BREAST_DF_RNC_perc_surv$`Death %` <- round(BREAST_DF_RNC_perc_surv$`Event Population`/BREAST_DF_RNC_perc_surv$Population*100,2)

kable(BREAST_DF_RNC_perc_surv)


#next let's look into the surgery and the survival rate and whether it migth have been critical or not. 



```

## Results of the exploratory data analysis

In this section, we look into some exploratory data analysis such as

-   Cause of death of those who have had cancer

-   Total number of tumors (Malignant or Benign)

-   Radiation and chemotherapy

-   Marital Status

We looked into the population and then among the population how many
survived the cancer. Later we will run some analyses to see whether
those were important or deciding factors or not.

```{r}

```
